Efficiency

### 2018-09-30 Second.Rnw
Final final, about to send

\usepackage{placeins}

@@

\usepackage{draftwatermark}
\SetWatermarkText{DRAFT}
\SetWatermarkScale{1.5}
\SetWatermarkLightness{0.95}

@@

\section{Executive Summary}
In January 2018, I produced a detailed report\parencite{morrison2018}, commissioned by Gary Johnston from Submarines for Australia, in order to examine the question of the efficiency of a pumpjet.   Pumpjets, whilst common on nuclear submarines which can operate at high speeds (above 25kt) for sustained periods, are generally not used on diesel-electric submarines, which need to operate at very low speeds (around 10kt during transit, and under 5kt during combat operations).  Considerable emphasis had been placed on the significance of the pumpjet proposed in the successful French bid made by DCNS (now Naval Group) for Australia's Future Submarine.  I concluded that an inevitable characteristic of pumpjets - directly necessitated by the characteristics which gave them good acoustic performance at high speeds - was very low propulsive performance (efficiency) at low speeds.  This would mean that a diesel-electric submarine would suffer poorer range and endurance, and significantly worse dived range and endurance, as well as a worse indiscretion ratio due to the use of a pumpjet.  Given the profound importance of these characteristics to all diesel-electric submarines, as well as the elevated importance of range and endurance to the Australian requirement, it would appear that a pumpjet can be safely regarded as entirely unsuitable for Australia's Future Submarine.

On June 7 2018, I appeared as a witness before the Senate Economic References Committee \parencite{SEN2018morrison}, where I defended and expanded upon my results.  My evidence was presented in the presence of senior officials from Defence and Naval Group, who steadfastly refused to comment.  The Committee directed Defence to provide a written response to my paper prior to the conclusion of the Inquiry into the Future of Australia's Naval Shipbuilding Industry (scheduled for the 27th of June) which they finally produced in mid September.  Submarines for Australia had already commissioned me to write a subsequent report in reply to the responses that had appeared in responses to Questions on Notice in Parliament, and also in the media.  With Defence's submission finally received, this paper is in a position to comprehensively assess the totality of the arguments put by Defence against the conclusions of my previous paper.

With the receipt of the written report it has become clear that Defence has no credible basis for rejecting the findings of my research.  Their arguments are so completely incongruous with the most essential and frankly obvious physical principles, it appears that Defences has quite thoroughly disgraced their technical credibility in making this response.

Their claim that pumpjets can be efficient across the entire speed range, including patrol speed, are instead based on the assumption that the efficiency of propulsors operating on fully-submerged bodies (like submarines) is ``essentially constant over speed"\parencite{defence2018} \parencite{kerr2018}.  This is supported by the argument that efficiency relates to resistance, and that resistance faced by underwater bodies does not vary with speed, since it is only wave-making resistance (encountered by vessels on the surface) that is the speed-dependent component of resistance.

@@

\begin{enumerate}
\item  \textbf{The French know the answer.}  As one of only two countries in the world who have designed and produced both pumpjets and propellers, nuclear and conventional submarines, and fitted and operated each type of propulsor on both types of submarines, there can be no doubt whatsoever that Naval Group and the French Government do have a well-formed opinion about the performance of pumpjets on diesel-electric submarines.  They would have extensive operational performance data from their in-service SSBNS which use the pumpjet.  They would have extremely detailed modelling, and tank-test results from the pumpjet from the almost-complete Barracuda class, which has an identical hull diameter to the proposed Shortfin Barracuda. They would have extensive operational performance data from their own Rubis Class nuclear submarines, which still use propellers, as well as the results of earlier tests, confirmed to have occurred by Naval Group Officials\parencite{SEN2018morrison}, where a pumpjet was fitted to a diesel-electric Agosta Class submarine.  That they have adequate performance data from all the necessary reference points to assess the likely performance of the proposed pumpjet on our own submarine is entirely beyond dispute.

@@

After the initial publication of my first report, three Questions on Notice were asked by Senator Ketter, specifically referencing my report[\ref{QON1}, \ref{QON2}, \ref{QON3}].  Whilst Defence avoiding answering all of them directly, one of them was particularly brazenly side-stepped.  Asked ``Is Defence still confident that Naval Group still hold the opinion that the pump-jet is well suited to diesel-electric submarines?" Defence merely reiterated their intent is to use the pumpjets as offered, strictly avoiding mentioning anyone's opinion, either their own or Naval Group's.  If the answer actually was what we would all assume it should be, that would have been a very easy question to answer directly.  A simple `yes' would be sufficient.  If the answer was `no', then we must assume that at the time of writing (presumably before the return date of 3 April, though Defence isn't always timely) that Defence had an inkling (at the very least) of what Naval Group undoubtedly knew all along - that pumpjets are entirely unsuitable for diesel-electric submarines.  In avoiding answering the question directly, Defence was not just misleading the Parliament, but clearly acting to shelter Naval Group from subsequent questions about fraudulent or misleading conduct during the CEP.

Giving Defence the most generous possible interpretation, one might allow the possibility that Defence were just coming to grips with being completely hood-winked during the CEP by their chosen design-partner, and perhaps felt that their wrath at being deliberately deceived ought not be publicly vented, or at least not at this stage.  One might, at a stretch, allow that maintaining the opportunity for the partner to save-face was still a useful bargaining-chip that our Defence chiefs to use in order to bring a potentially treacherous contractor to heel. (Presumeably in addition to a range of contractual penalties, as well as the commercial leverage the Commonwealth would have by being able to re-introduce competition into the design process as many have suggested ought be done in any case \parencite{stanford2017}, or simply not proceed to a build contract on completion of the design.)

The state of commercial negotiations between the Commonwealth and Naval Group, as revealed in the Senate Economic References Committee \parencite{SEN2018morrison}, show that Defence was moving in exactly the opposite direction - shocking Senators with the revelation that Defence was looking instead to contract immediately with DCNS for the full construction of all 12 submarines, to the complete exclusion of the Australian Submarine Corporation, within coming months as part of the Strategic Partnering Agreement.  This would effectively wipe out the opportunity for any competitive tension at any stage of the \$79 billion five-decade project within the first two years of a five-year design phase.

@@

In summary, I conclude that the pumpjet was a fraudulent, rather than incomptent proposal, used to construct the fiction that France's vaunted heritage as a designer of nuclear submarines offers a substantial technical advantage to a non-nuclear submarine.  There is almost no chance that it will be included in the detailed design.  In reality, without a nuclear reactor, there is relatively little relevant nuclear-related technology or experience (other than the combat system, mandated from a 3rd party) which is usefully transferable.  The extent to which the bottom of the barrel needed to be scraped to maintain this fiction was illustrated when RADM Sammut was challenged on the benefit of the nuclear design after facing pressure on the pumpjet, and was forced to refer to the `diameter of the hull' as being a key learning[\ref{SENhanson}].  Given the relatively more accessible engineering concepts involved, such claim would be much more easily subjected to public ridicule had it been more substantially promoted, as the pumpjet was.\footnote{Among the many engineering challenges presented in designing a new submarine, ensuring the structural integrity of the hull at a new diameter is one of the most well-understood and predictable.  The Collins Class was evolved from a submarine of smaller hull diameter, and this aspect of the design and development encountered no problems.  In 2006, DSTO commissioned a paper suggesting that the Collins Class replacement be designed at a wider hull, which didn't correspond to that of the Barracuda.\parencite{joubert2006}}  In contrast, those countries whose own navies relied entirely upon diesel-electric submarines would likely boast significant advantages in technologies which are genuine game-changers for diesel-electric submarines, such as advanced composite propellers, Air-Independent Propulsion, and Lithium-Ion batteries, which we are now told won't be included in the Future Submarine \textit{because} of the pumpjet\parencite{bergmann2018}.  To that extent, it is almost certain that the pumpjet proposal will not only impose a substantial cost on Australian taxpayers (tens of billions of dollars), but also substantially degrade Australia's future military capability.

In light of these considerations, Defence should be required to explain again what their reasons were for choosing the French as a design partner, as there can now be no doubt that the pumpjet played no legitimate part in it.  Any claims to the contrary are made either out of incompetence, as Defence's latest response might have us believe, or are made with the deliberate intent of misleading.

@@

In early September, the Department of Defence gave a six-page written response to my paper, a full three months after the initial request. The two substantive arguments already identified were repeated, but the weight of their argument actually came to rest on another assumption.  This assumption was mentioned already by Dr Kershaw in the ADM article, but it was so completely at odds with the most basic physics that I was inclined to dismiss it as perhaps a clumsily worded statement made in an interview compounded by a little journalistic mis-quotation, rather than a substantive argument.   The re-appearance of this argument, which was spelled out and repeated multiple times in a written document that was prepared by Defence over the course of three months (well after any nominal deadline for timeliness had elapsed) dispelled all illusions that the argument was included by accident.  As such, I have also addressed it here as one of three substantial arguments made against my earlier conclusions regarding pumpjet efficiency.  This third argument coming from a document so cafeully and deliberately prepared has had a material impact on the strength of the conclusions of this paper, which seeks to consider all three lines of reasoning in sufficient detail to suitably discuss the circumstances in which such arguments might be made.

@@

\newpage
\section{A brief chronology of comment}

@@

\subsection{Department of Defence, to Senator Ketter's Questions on Notice}
\begin{tcolorbox}[width=0.9\textwidth, center, colback=black!5!white,colframe=black!75!black, title= \textbf{3 April 2018 - Department of Defence, in response to Question on Notice from Senator Chris Ketter in December 2018}]
\textit{Senator Ketter asks Question 1, referencing specifically Greg Sammut in \parencite{sammut2017} and my paper \parencite{morrison2018} asks specifically whether Defence maintains that the pumpjet can be efficient across the entire speed range. Full question can be found in \ref{QON1}.}
\\
\\
Question: Does defence still maintain that a pump-jet can be efficient across the entire speed range?
\\
\\
Answer: ``The efficiency of pump-jets designed for submarine use cannot be determined through comparisons with water-jets and ducted propellers, which are different in their operation from pump-jets. Moreover, the global efficiency of submarine pump-jets cannot be derived from `open water' tests of pump-jets, where the pump-jet is assessed independently of the hull of the submarine. A correctly designed pump-jet, suited to the aft hull shape of the submarine, achieves better optimisation of efficiency and stealth than any propeller, even at low speeds."

@@

\begin{tcolorbox}[width=0.9\textwidth, center, colback=black!5!white,colframe=black!75!black, title= \textbf{3 April 2018 - Department of Defence, in response to Question on Notice from Senator Chris Ketter in December 2018}]
\textit{Senator Ketter asks Question 2, inquiring as to how Defence assesses Naval Group's opinion regarding the pumpjet's suitability, following Jean-Michel Billig's comments.  Full questions and responses  can be found in \ref{QON2}.}
\\
\\
Question: Is Defence still confident that Naval Group still hold the opinion that the pump-jet is well suited to diesel-electric submarines?
\\
\\
Answer: ``Defence intends to use the pump jet propulsion system for the Future Submarine that was offered by Naval Group as part of their Competitive Evaluation Proposal and remains on offer to Australia."

@@

@@

\begin{tcolorbox}[width=0.9\textwidth, center, colback=green!5!white,colframe=green!75!black, title= \textbf{How a scientist feels about being told that because underwater bodies don't make waves, the resistance they encounter is constant with speed}]

@@

\begin{figure}[h]
\captionsetup{width=0.7\textwidth}
\includegraphics[width=0.7\textwidth, center]{drag4.png}
\caption{A breakdown of the components of resistance taken from Reference \parencite{mollard2011} does indeed reveal that deeply submerged submarines encounter no wave-making pressure, but still encounter frictional resistance (which scales with velocity squared).}
\label{fig:drag4.png}
\end{figure}

\begin{figure}[h]
\captionsetup{width=0.7\textwidth}
\includegraphics[width=0.7\textwidth, center]{dragfroude.png}
\caption{A table showing the results of Froude's original experiments used to measure skin friction of various surfaces in the 1870s, taken from reference \parencite{lewis1988}.  Note that the exponent of velocity is determined consistently to be about 2. }
\label{fig:dragfroude.png}
\end{figure}

\begin{figure}[h]
\captionsetup{width=0.7\textwidth}
\includegraphics[width=0.7\textwidth, center]{drag1.png}
\caption{Taken from the appendix of \underline(Fundamentals of Submarine Concept  Design) by  CAPT Harry A. Jackson USN (Ret.) \parencite{jackson1992}.  Note that the $V^2$ which indicates the relationship with velocity squared relationship}
\label{fig:drag1.png}
\end{figure}

\begin{figure}[h]
\captionsetup{width=0.7\textwidth}
\includegraphics[width=0.7\textwidth, center]{drag3.png}
\caption{Taken from Research Paper commissioned by Department of Navy (US) titled \underline{Elements of the drag of underwater bodies} \parencite{granville1976}.  Note that the $U^2$ which indicates the relationship with velocity squared relationship}
\label{fig:drag3.png}
\end{figure}

\begin{figure}[h]
\captionsetup{width=0.7\textwidth}
\includegraphics[width=0.7\textwidth, center]{drag2.png}
\caption{Taken from conference paper by Professor Volker Bertram titled \underline{Submarine Hull Design} \parencite{bertram2012}, same equation can also be found in published textbook by same author \parencite{bertram2000}.  Note that the $V^2$ which indicates the relationship with velocity squared relationship}
\label{fig:drag2.png}
\end{figure}

@@

\begin{tcolorbox}[width=0.9\textwidth, center, colback=green!5!white,colframe=green!75!black, title= \textbf{How a scientist feels being told that the pumpjet is a game-changer for our diesel-electric submarine because it is so large, where it wouldn't make sense on a smaller one}]

@@

\begin{tcolorbox}[width=0.9\textwidth, center, colback=green!5!white,colframe=green!75!black, title= \textbf{How a scientist feels about being told the pumpjet efficiency can't be assessed because it depends on the hull}]

@@

@@

@@

@@

\end{appendices}

>>>>

\usepackage{placeins}
\usepackage{wrapfig}

@@

% \usepackage{draftwatermark}
% \SetWatermarkText{DRAFT}
% \SetWatermarkScale{1.5}
% \SetWatermarkLightness{0.95}

@@

\section{Executive Summary}
In January 2018, I produced a detailed report\parencite{morrison2018}, commissioned by Gary Johnston from Submarines for Australia, in order to examine the question of the efficiency of a pumpjet.   Pumpjets, whilst common on nuclear submarines which can operate at high speeds (above 25kt) for sustained periods, are generally not used on diesel-electric submarines, which need to operate at very low speeds (around 10kt during transit, and under 5kt during combat operations).  Considerable emphasis had been placed on the significance of the pumpjet proposed in the successful French bid made by DCNS (now Naval Group) for Australia's Future Submarine.  I concluded that an inevitable characteristic of pumpjets - directly necessitated by the characteristics which gave them good acoustic performance at high speeds - was very low propulsive performance (efficiency) at low speeds.  This would mean that a diesel-electric submarine would suffer poorer range and endurance, and significantly worse dived range and endurance, as well as a worse indiscretion ratio due to the use of a pumpjet.  Given the profound importance of these characteristics to all diesel-electric submarines, as well as the elevated importance of range and endurance to the Australian requirement, it appears from these results that a pumpjet can be safely regarded as entirely unsuitable for Australia's Future Submarine.

On June 7 2018, I appeared as a witness before the Senate Economic References Committee \parencite{SEN2018morrison}, where I defended and expanded upon my results.  My evidence was presented in the presence of senior officials from Defence and Naval Group, who steadfastly refused to comment.  The Committee directed Defence to provide a written response to my paper prior to the conclusion of the Inquiry into the Future of Australia's Naval Shipbuilding Industry (scheduled for the 27th of June) which they finally produced in mid September.  Submarines for Australia had already commissioned me to write a subsequent report in reply to the responses that had appeared in answers to Questions on Notice in Parliament, and also in the media.  With Defence's submission finally received, this paper is in a position to comprehensively assess the totality of the arguments put by Defence against the conclusions of my previous paper.

With the receipt of the written report it has become clear that Defence has no credible basis for rejecting the findings of my research.  Their arguments are so completely incongruous with the most essential (and relatively obvious) physical principles, it appears that Defences has quite thoroughly disgraced their technical credibility in making this response.

Their claim that pumpjets can be efficient across the entire speed range, including patrol speed, are based on the assumption that the efficiency of propulsors operating on fully-submerged bodies (like submarines) is ``essentially constant over speed"\parencite{defence2018} \parencite{kerr2018}.  This is supported by the argument that efficiency relates to resistance, and that resistance faced by underwater bodies does not vary with speed, since it is only wave-making resistance (encountered by vessels on the surface) that is the speed-dependent component of resistance.

@@

\begin{enumerate}
\item  \textbf{The French know the answer.}  As one of only two countries in the world who have designed and produced both pumpjets and propellers, nuclear and conventional submarines, and fitted and operated each type of propulsor on both types of submarines, there can be no doubt whatsoever that Naval Group and the French Government do have a well-formed opinion about the performance of pumpjets on diesel-electric submarines.  They would have extensive operational performance data from their in-service SSBNs which use the pumpjet.  They would have extremely detailed modelling, and tank-test results from the pumpjet from the almost-complete Barracuda class, which has an identical hull diameter to the proposed Shortfin Barracuda. They would have extensive operational performance data from their own Rubis Class nuclear submarines, which still use propellers, as well as the results of earlier tests, confirmed to have occurred by Naval Group Officials\parencite{SEN2018morrison}, where a pumpjet was fitted to a diesel-electric Agosta Class submarine.  That they have adequate performance data from all the necessary reference points to assess the likely performance of the proposed pumpjet on our own submarine is entirely beyond dispute.

@@

After the initial publication of my first report, three Questions on Notice were asked by Senator Ketter, specifically referencing my report[\ref{QON1}, \ref{QON2}, \ref{QON3}].  Whilst Defence avoided answering all of them directly, one of them was particularly brazenly side-stepped.  Asked ``Is Defence still confident that Naval Group still hold the opinion that the pump-jet is well suited to diesel-electric submarines?" Defence merely reiterated their intent is to use the pumpjets as offered, strictly avoiding mentioning anyone's opinion, either their own or Naval Group's.  If the answer actually was what we would all assume it should be, that would have been a very easy question to answer directly.  A simple `yes' would be sufficient.  If the answer was `no', then we must assume that at the time of writing (presumably before the return date of 3 April, though Defence isn't always timely) that Defence had an inkling (at the very least) of what Naval Group undoubtedly knew all along - that pumpjets are entirely unsuitable for diesel-electric submarines.  In avoiding answering the question directly, Defence was not just misleading the Parliament, but clearly acting to shelter Naval Group from subsequent questions about fraudulent or misleading conduct during the CEP.

Giving Defence the most generous possible interpretation, one might allow the possibility that Defence was just coming to grips with being completely hood-winked during the CEP by their chosen design-partner, and perhaps felt that their wrath at being deliberately deceived ought not be publicly vented, or at least not at this stage.  News that negotiations have largely broken-down\parencite{greene2018} could well be considered in this light.  One might, at a stretch, allow that maintaining the opportunity for the partner to save-face was still a useful bargaining-chip that our Defence chiefs to use in order to bring a potentially treacherous contractor to heel. (Presumably in addition to a range of contractual penalties, as well as the commercial leverage the Commonwealth would have by being able to re-introduce competition into the design process as many have suggested ought be done in any case \parencite{stanford2017}, or simply not proceed to a build contract on completion of the design.)

However, at least at the time of the final public hearing for the Inquiry into the Future of Naval Shipbuilding in Australia \parencite{SEN2018morrison}, it appeared that Defence was moving in exactly the opposite direction.  Senators on the committee were deeply concerned with establishing just how early it was during or after the CEP process that Defence agreed (with Naval Group) that they would be the \textit{builder} of the submarines, not just the design partner as previously assumed, and that Australian Submarine Corporation was to be consumed by DCNS.\footnote{On the 20th of June 2017 the Acting CEO of DCNS, Brent Clark stated that ``from our perspective, ASC will be consumed by DCNS."}  A further topic of substantial interest was whether the Strategic Partnering Agreement (SPA) that was currently being negotiated would commit the Commonwealth to use DCNS as the \textit{builder} of \textit{all 12} submarines. It was clear to anyone in the room that Defence's responses implied that it was very likely that this would be the case.

Taken together, these observations paint a very different picture of the intended direction of Defence's relationship with DCNS.  Defence appeared to be excluding two options which might have been assumed as likely or even default courses of action for a government seeking to minimise risk and retain some sovereign control over a crucial defence project:  either that the chosen foreign design partner would be expected to work with a local builder (one would assume ASC) to deliver the build, or that allocation of the actual build contract (or contracts, in the case they proceed in blocks) be left until a later stage when the detailed design (or preceding block of submarines) was closer to being completed satisfactorily.  As members of the committee were clearly conscious, the apparent course taken by defence would effectively wipe out the opportunity for any competitive tension (or external technical scrutiny) at any stage of the \$79 billion five-decade project within the first two years of a five-year detailed design phase.  Why Defence might be contemplating eliminating, rather than exercising opportunities to increase competitive tension at a time when it was becoming clear that the pumpjet proposal was disingenuous is extremely unclear.

@@

In summary, I conclude that the pumpjet was a fraudulent, rather than incompetent proposal, used to construct the fiction that France's heritage as a designer of nuclear submarines offers a substantial technical advantage to a non-nuclear submarine.   In reality, without a nuclear reactor, there is essentially no nuclear-related technology or experience which is usefully transferable to a diesel-electric submarine.  The extent to which the bottom of the barrel needed to be scraped to maintain this fiction was illustrated when RADM Sammut was challenged on the benefit of the nuclear design after facing pressure on the pumpjet, and was forced to refer to the `diameter of the hull' as being a key learning[\ref{SENhanson}].  Given the relatively more accessible engineering concepts involved, such a claim would be much more easily subjected to public ridicule had it been more substantially promoted, as the pumpjet was.\footnote{Among the many engineering challenges presented in designing a new submarine, ensuring the structural integrity of the hull at a new diameter is one of the most well-understood and predictable.  The Collins Class was evolved from a submarine of smaller hull diameter, and this aspect of the design and development encountered no problems.  In 2006, DSTO commissioned a paper suggesting that the Collins Class replacement be designed at a wider hull, which didn't correspond to that of the Barracuda.\parencite{joubert2006}}  In contrast, those countries whose own navies relied entirely upon diesel-electric submarines would likely boast significant advantages in technologies which are genuine game-changers for diesel-electric submarines, such as advanced composite propellers, Air-Independent Propulsion, and Lithium-Ion batteries, which we are now told won't be included in the Future Submarine \textit{because} of the pumpjet\parencite{bergmann2018}.  To that extent, it is almost certain that the pumpjet proposal will not only impose a substantial cost on Australian taxpayers (tens of billions of dollars), but also substantially degrade Australia's future military capability.

In light of these considerations, Defence should be required to explain again what their reasons were for choosing the French as a design partner, as there can now be no doubt that the pumpjet played no legitimate part in it.  The use of a pumpjet will inevitably cause a diesel-electric submarine to have poorer endurance and range, particularly dived endurance and range, and a poorer indiscretion ratio than one with open-tipped propellers.  Any claims to the contrary are made either out of incompetence, as Defence's latest response might have us believe, or are made with the deliberate intent of misleading.

@@

In early September, the Department of Defence gave a six-page written response to my paper, a full three months after the initial request. The two substantive arguments already identified were repeated, but the weight of their argument actually came to rest on another assumption.  This assumption was mentioned already by Dr Kershaw in the ADM article, but it was so completely at odds with the most basic physics that I was inclined to dismiss it as perhaps a clumsily worded statement made in an interview compounded by a little journalistic mis-quotation, rather than a substantive argument.   The re-appearance of this argument, which was spelled out and repeated multiple times in a written document that was prepared by Defence over the course of three months (well after any nominal deadline for timeliness had elapsed) dispelled all illusions that the argument was included by accident.  As such, I have also addressed it here as one of three substantial arguments made against my earlier conclusions regarding pumpjet efficiency.  This third argument coming from a document so carefully and deliberately prepared has had a material impact on the strength of the conclusions of this paper, which seeks to consider all three lines of reasoning in sufficient detail to suitably discuss the circumstances in which such arguments might be made.

@@

\newpage
\section{A chronology of comment}

@@

\subsection{Department of Defence, to Senator Ketter's Questions on Notice}
\begin{tcolorbox}[width=0.9\textwidth, center, colback=black!5!white,colframe=black!75!black, title= \textbf{3 April 2018 - Department of Defence, in response to Question on Notice from Senator Chris Ketter with return date of 3 April 2018}]
\textit{Senator Ketter asks Question 1, referencing specifically Greg Sammut in \parencite{sammut2017} and my paper \parencite{morrison2018} asks specifically whether Defence maintains that the pumpjet can be efficient across the entire speed range. Full question can be found in \ref{QON1}.}
\\
\\
Question: Does defence still maintain that a pump-jet can be efficient across the entire speed range?
\\
\\
Answer: ``The efficiency of pump-jets designed for submarine use cannot be determined through comparisons with water-jets and ducted propellers, which are different in their operation from pump-jets. Moreover, the global efficiency of submarine pump-jets cannot be derived from `open water' tests of pump-jets, where the pump-jet is assessed independently of the hull of the submarine. A correctly designed pump-jet, suited to the aft hull shape of the submarine, achieves better optimisation of efficiency and stealth than any propeller, even at low speeds."

@@

\begin{tcolorbox}[width=0.9\textwidth, center, colback=black!5!white,colframe=black!75!black, title= \textbf{3 April 2018 - Department of Defence, in response to Question on Notice from Senator Chris Ketter with return date 3 April 2018}]
\textit{Senator Ketter asks Question 2, inquiring as to how Defence assesses Naval Group's opinion regarding the pumpjet's suitability, following Jean-Michel Billig's comments.  Full questions and responses  can be found in \ref{QON2}.}
\\
\\
Question: Is Defence still confident that Naval Group still hold the opinion that the pump-jet is well suited to diesel-electric submarines?
\\
\\
Answer: ``Defence intends to use the pump jet propulsion system for the Future Submarine that was offered by Naval Group as part of their Competitive Evaluation Proposal and remains on offer to Australia."

@@

\subsection{Department of Defence, to Senator Patrick's Questions on Notice}
\begin{tcolorbox}[width=0.9\textwidth, center, colback=black!5!white,colframe=black!75!black, title= \textbf{Sept 2018 - Department of Defence, in response to Question on Notice from Senator Rex Patrick with return date 18 July 2018}]
\textit{Senator Patrick asks a series of questions regarding the technical readiness level (TRL) of various technologies in the Future Submarine Program, and which ones were included or excluded in the French pre-concept design in the CEP. Full questions and answers can be found in Appendix \ref{QON827}.}
\\
\\
\textbf{Question:} What is the Technological Readiness Level (TRL) of the propulsor being advanced for inclusion in the future submarine design.

\textbf{Answer:} The technology to be used in the propulsor for the Future Submarine is assessed to be at TRL 9 noting it has been in operational use since the end of the 1990s on French submarines. The propulsor for the Future Submarine will be based on this technology.
\\
\\
\textbf{Question:} Has an AIP been ruled out as a Future Submarine System in the early design work.

\textbf{Answer:} At this point of the design process, increased battery and recharging capacity is proving more effective than the addition of AIP in meeting Australia's submarine mission requirements.

This response to Questions on Notice clearly demonstrates that the technology being referred to for the pumpjet is indeed the in-service design from the Triomphant Class of SSBNs in the French Navy.  There is no conceivable way that any technology at a higher level of readiness than TRL 7 would not have detailed performance data available, across the full range of operating conditions\parencite{DSTG2018a}.  Alternatively, if the proposal is actually to design a new pumpjet ground-up for the radically different operating speed profile, then this answer is almost certainly misleading, as the actual propulsor to be used is probably at TRL 2, or at best TRL 3.  However, as I have argued before the Senate[\ref{SENmorrison}] and in Reference \parencite{morrison2018}, any attempt to develop a pumpjet that is as efficient as a propeller at low speeds is essentially impossible. The elevated pressure (and lower speed) of water at the rotor which allows the superior acoustic performance at high speed \textit{necessitates} the decline in efficiency at low speeds.

The comment regarding AIP also confirms that RADM Sammut's words in the APDR are not simply careless statements to the media or poor reporting, but represent a considered position of Defence - that AIP is still being considered, (i.e. not ruled out) but doesn't currently appear superior in meeting `mission requirements'.  Another clear implication of this statement is that the mission requirements do exist, and are fixed.  If they are not so, then this statement would be most severely misleading.  This response also seems to confirm that a back-to-front design process is indeed in place, whereby the size of the hull is fixed, whilst there is still a large enough portion of the available displacement unallocated that one could consider an AIP system (several hundreds of tonnes) as one of a plurality of options to include.  That Defence requires an \textit{increased} level of batteries in order to meet the requirement can only mean that the previous nominal load of batteries was insufficienct to meet the requirement, or that the plan now is to very substantially exceed the requirement.

As was discussed earlier in this section, the scale of the impact in terms of performance and displacement of major technology choices such as the type of batteries, type of propulsor, and inclusion of AIP is so large that their status would need to be determined very early in pre-concept design, well in advance of the final hull dimensions being finalised.  Otherwise the resultant variation in displacement/performance will result either in a potential engineering catastrophe (of Super-Seasprite proportions, i.e. the vessel doesn't sink and/or float), or a performance outcome that will not even remotely resemble the performance specified in the requirement.

@@

\begin{tcolorbox}[width=0.9\textwidth, center, colback=olive!5!white,colframe=olive!75!black, title= \textbf{How a scientist feels about being told that because underwater bodies don't make waves, the resistance they encounter is constant with speed}]

@@

\begin{figure}[h]
\captionsetup{width=0.9\textwidth}
\includegraphics[width=0.7\textwidth, center]{drag4.png}
\caption{A breakdown of the components of resistance taken from Reference \parencite{mollard2011} does indeed reveal that deeply submerged submarines encounter no wave-making pressure, but still encounter frictional resistance (which scales with velocity squared).}
\label{fig:drag4.png}
\end{figure}

\begin{figure}[h]
\captionsetup{width=0.9\textwidth}
\includegraphics[width=0.7\textwidth, center]{dragfroude.png}
\caption{A table showing the results of Froude's original experiments used to measure skin friction of various surfaces in the 1870s, taken from reference \parencite{lewis1988}.  Note that the exponent of velocity is determined consistently to be about 2. }
\label{fig:dragfroude.png}
\end{figure}

\begin{figure}[h]
\captionsetup{width=0.9\textwidth}
\includegraphics[width=0.7\textwidth, center]{drag1.png}
\caption{Taken from the appendix of \underline(Fundamentals of Submarine Concept  Design) by  CAPT Harry A. Jackson USN (Ret.) \parencite{jackson1992}.  Note that the $V^2$ which indicates the relationship with velocity squared relationship}
\label{fig:drag1.png}
\end{figure}

\begin{figure}[h]
\captionsetup{width=0.9\textwidth}
\includegraphics[width=0.7\textwidth, center]{drag3.png}
\caption{Taken from Research Paper commissioned by Department of Navy (US) titled \underline{Elements of the drag of underwater bodies} \parencite{granville1976}.  Note that the $U^2$ which indicates the relationship with velocity squared relationship}
\label{fig:drag3.png}
\end{figure}

\begin{figure}[h]
\captionsetup{width=0.9\textwidth}
\includegraphics[width=0.7\textwidth, center]{drag2.png}
\caption{Taken from conference paper by Professor Volker Bertram titled \underline{Submarine Hull Design} \parencite{bertram2012}, same equation can also be found in published textbook by same author \parencite{bertram2000}.  Note that the $V^2$ which indicates the relationship with velocity squared relationship}
\label{fig:drag2.png}
\end{figure}

@@

\begin{tcolorbox}[width=0.9\textwidth, center, colback=olive!5!white,colframe=olive!75!black, title= \textbf{How a scientist feels being told that the pumpjet is a game-changer for our diesel-electric submarine because it is so large, where it wouldn't make sense on a smaller one}]

@@

\begin{tcolorbox}[width=0.9\textwidth, center, colback=olive!5!white,colframe=olive!75!black, title= \textbf{How a scientist feels about being told the pumpjet efficiency can't be assessed because it depends on the hull}]

@@

\section{About the Author}

\begin{wrapfigure}{r}{0.35\textwidth}
  \begin{center}
    \includegraphics[width=0.28\textwidth]{aidan.jpg}
  \end{center}
  \caption{Aidan Morrison}
\end{wrapfigure}

Aidan Morrison is a physicist, entrepreneur and data-scientist who lives in Sydney with his wife Constance.  He graduated from the Australian National University with First Class Honours in Theoretical Particle Physics, and commenced a PhD in Experimental Particle Physics, doing work on the ATLAS Experiment at CERN searching for the Higgs Boson.

He left his PhD incomplete in order to found Rubber Ducky Defence, a start-up dedicated to developing small naval platforms and weapons for swarm warfare, and chose waterjets and pumpjets respectively for the  semi-submersible Swarm Combat Vessel and medium-weight torpedo involved.  Rubber Ducky responded to Requests for Proposal (RFPs) for SEA 1778 and JP 1770, by adapting this vessel to meet a requirement for deployable mine-countermeasures and rapid environmental assessment.  This requirement included towing and deploying a range of fully-submerged sensors, decoys, remotely and autonomously operated underwater vehicles in a variety of speeds to meet the specified conceptual operations.  Endurance and range were crucial considerations in responding to the Preliminary Function and Performance Specifications for these two projects.  A detailed proposal evolved from this concept and submitted to DSTO's Concept Technology Demonstrator program in 2013 was acknowledged as ``probably a world-class development".

His other accomplishments include co-founding a bar-restaurant in Randwick called Bat Country, and developing a predictive forecasting tool for hospitality businesses using machine-learning techniques under Trendlock, which also undertakes additional consulting services.  He has also lead a team of economic analysts at boutique think-tank, and currently works as a data-scientist at Tridant, most recently using machine-learning and statistical techniques to implement predictive maintenance routines for heavy industrial machinery.

@@

\begin{flushleft}

@@

\end{flushleft}

@@

\subsection{Answer to Question on Notice}\label{QON827}

\begin{figure}[h]
\includegraphics[width=0.7\textwidth, center]{QON827a.png}
\end{figure}

\begin{figure}[h]
\includegraphics[width=0.7\textwidth, center]{QON827b.png}
\end{figure}

\FloatBarrier
\end{appendices}
